---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
- Load data into dataframe "activityData" using `read.csv()` 
- Remove `NA`
- Convert dates

```{r}
input<-read.csv("activity.csv")
activityData<-input[complete.cases(input),]
head(activityData)
```


Aggregate data into daily steps
```{r}
activityDays<-aggregate(activityData[, 1], list(activityData$date), sum)
names(activityDays)<-c("Date","Steps")
```

## What is mean total number of steps taken per day?
Create histogram
```{r, fig.height=4}
hist(activityDays$Steps, col = "red",xlab="Steps daily",main="",breaks=15)
```


Calculate simple statistics:
```{r}
mean(activityDays$Steps)
median(activityDays$Steps)
```

So, the mean total number of steps is very close to median and equal 10766.

## What is the average daily activity pattern?
Again convert data:
```{r}
activityIntervals<-aggregate(activityData[, 1], list(activityData$interval), mean)
plot(activityIntervals$x,type="l",xlab="5-min intervals",ylab="average steps per interval")
```

Maximum steps in this 5-minute interval:
```{r}
activityIntervals[activityIntervals$x==max(activityIntervals$x),]$Group.1
```
8:35 could means morning exercises of walk to office by foot.

## Imputing missing values
Calculate number of NA:
```{r}
sum(is.na(input$steps))
```
It's about
```{r}
round(sum(is.na(input$steps))/nrow(input),2)

```
 of data.
 
 Function that enriches data with average daily steps for current interval:
```{r}
newData<-input
fg<-function(){
  for (i in 1:nrow(newData)){
    if(is.na(newData$steps[i])==TRUE){
      newData$steps[i]<-activityIntervals[activityIntervals$Group.1==newData$interval[i],2]
    } 
  }
  newData
}
newData<-fg()

```
 Let's check on `NA` again:
```{r}
 sum(is.na(newData$steps))
```

```{r}
activityDays2<-aggregate(newData[, 1], list(newData$date), sum)
```

Create histogram:
```{r, fig.height=4}
hist(activityDays2$x, col = "blue",xlab="Steps daily",main="Without NA",breaks=15)
```

Calculate simple statistics:
```{r}
mean(activityDays2$x)
median(activityDays2$x)
```
When we add some non-zero values into data we increase median.


## Are there differences in activity patterns between weekdays and weekends?

I didn't complete this part, because a met an error in script below and I had not enough time to solve it.

Create factors
newData$date=as.Date(newData$date,"%y-%m-%d")
newData$WD <- ifelse(c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")[as.POSIXlt(newData$date)$wday + 1]=="Saturday"|c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday","Friday", "Saturday")[as.POSIXlt(newData$date)$wday + 1]=="Sunday", "weekend", "weekday")

Subset data into two factors
newData1<-subset(newData,newData$WD=="weekday")
activityIntervals1<-aggregate(newData1[, 1], list(newData1$interval), mean)
newData2<-subset(newData,newData$WD=="weekend")
activityIntervals2<-aggregate(newData2[, 1], list(newData2$interval), mean)

Plot results
plot(activityIntervals1$x, type="n",xlab="",main="",
     ylab="average steps per interval" ,xaxt='n')
plot(activityIntervals1$x,type="l",col="red")
plot(activityIntervals2$x,type="l",col="blue")
legend("topright",legend=c("Weekdays","Weekends"),
       lty=1,col=c("red","blue"),y.intersp=0.3,bty="n")

